---
title:           Covid-19 
author:          Dennis Wollersheim 
date:            06.03.2020
fig_width:       8 
fig_height:      4 
linkcolor:       cyan
citecolor:       grey
output:
  pdf_document:
    highlight:   zenburn
---

\tableofcontents

```{r loadd, results='asis', echo=FALSE, include=FALSE,}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

devtools::install_github("RamiKrispin/coronavirus")
#install.packages('wpp2017')



```

# Data Source

Sourced from https://ramikrispin.github.io/coronavirus/ and the R package wpp2017, world population profile

If you want to know why Italy just quarantined 18 million people, read on.


```{r Setup}

suppressMessages(suppressWarnings( library(tidyverse)))
library(coronavirus)
library(wpp2017)
library(lubridate)
library(dplyr)
data(pop)

pop %>% 
  as_tibble() %>% 
  { . } -> pop

mapping = tribble(
                  ~Country.Region, ~name, 
                  'US', 'United States of America',
                  'United Kingdom', 'United Kingdom',
                  'China', 'China',
                  'Korea, South', 'Republic of Korea',
                  'Taiwan*','China, Taiwan Province of China',
                  'Vietnam', 'Viet Nam',
                  'Russia', 'Russian Federation',
                  'Iran','Iran (Islamic Republic of)'
)

coronavirus %>%
  as_tibble() %>%
  left_join( mapping ) %>% 
  mutate( name = ifelse(is.na(name), Country.Region, name)) %>%
  { . } -> coronavirus

if (FALSE) {
  mapping %>% 
    anti_join( coronavirus)

  mapping %>% 
    anti_join( pop )

  coronavirus %>%
    left_join( mapping ) %>% 
    as_tibble() %>%
    distinct(name) %>% 
    { . } -> ac
      pop %>% 
        filter( !str_detect(as.character(name), 'WORLD|countries|regions') ) %>%
        distinct(name) %>% anti_join(ac)
}

coronavirus %>%
  filter( type=='confirmed' ) %>%
  group_by( Country.Region ) %>%
  summarise( cases = sum(cases)) %>%
  top_n(10, cases) %>% 
  pluck('Country.Region') %>% 
  c('Australia', 'Singapore') %>%
  { . } -> top_10_countries

```
# Cumulative cases by country and number of days since first case

```{r}


coronavirus %>% 
  filter( type=='confirmed' ) %>%
  group_by( Country.Region ) %>%
  mutate(ndays = date-min(date) ) %>%
  group_by( Country.Region, ndays ) %>%
  summarise( cases=sum(cases)) %>%
  ungroup() %>%
  group_by( Country.Region) %>%
  filter( Country.Region %in% top_10_countries) %>%
  arrange( ndays, .by_group=TRUE) %>%
  mutate( cum_cases = cumsum( cases) ) %>%
  ggplot(aes( ndays, cum_cases, color=Country.Region)) +
  geom_line() +
  scale_y_log10() +
  scale_x_continuous()


```


# Cumulative deaths by country and number of days since first case

```{r}

coronavirus %>% 
  filter( type=='death' ) %>%
  group_by( Country.Region ) %>%
  mutate(ndays = date-min(date) ) %>%
  group_by( Country.Region, ndays ) %>%
  summarise( cases=sum(cases)) %>%
  ungroup() %>%
  group_by( Country.Region) %>%
  filter( Country.Region %in% top_10_countries) %>%
  arrange( ndays, .by_group=TRUE) %>%
  mutate( cum_cases = cumsum( cases) ) %>%
  ggplot(aes( ndays, cum_cases, color=Country.Region)) +
  geom_line() + 
  scale_y_log10() +
  scale_x_continuous()




```

# Rate of new cases per day, since the start of infection by country

 - public health works
 - individualism not so much
 - the hockeys stick curve is worrying people


```{r}

coronavirus %>% 
  filter( type=='confirmed' ) %>%
  filter( Country.Region != 'Others') %>%
  group_by( Country.Region ) %>%
  mutate(ndays = date-min(date) ) %>%
  group_by( Country.Region, ndays ) %>%
  summarise( cases=sum(cases)) %>%
  ungroup() %>%
  group_by( Country.Region) %>%
  filter( Country.Region %in% top_10_countries) %>%
  arrange( ndays, .by_group=TRUE) %>%
  mutate( rate = cumsum(cases ) / as.integer(ndays)  ) %>%
  ggplot(aes( ndays, rate, color=Country.Region)) +
  geom_smooth() +
  scale_y_log10() +
  scale_x_continuous()

```



# Total cumulative confirmed cases as a population rate
 - greater than 200 cases, or Singapore or Australia

Note, Japan had -1 cases on day 2

```{r}

coronavirus %>% 
  inner_join(pop, by='name' ) %>% 
  filter( type=='confirmed' ) %>%
  group_by( Country.Region ) %>%
  mutate(ndays = date-min(date) ) %>%
  group_by( Country.Region, ndays ) %>%
  summarise( cases=sum(cases), population = min(`2015`)) %>%
  ungroup() %>%
  group_by( Country.Region) %>%
  filter( Country.Region %in% top_10_countries) %>%
  arrange( ndays, .by_group=TRUE) %>%
  mutate( cumcases = cumsum(cases), rate = cumsum(cases ) / population  ) %>%
  ggplot(aes( ndays, rate, color=Country.Region)) +
  geom_line() +
  scale_y_log10() +
  scale_x_continuous()


```



# New cases population rate by week


```{r}

coronavirus %>% 
  inner_join(pop, by='name' ) %>% 
  filter( type=='confirmed' ) %>%
  group_by( Country.Region ) %>%
  mutate(week = week(date) ) %>%
  group_by( Country.Region, week ) %>%
  summarise( cases=sum(cases), population = min(`2015`)) %>%
  ungroup() %>%
  group_by( Country.Region) %>%
  # filter( Country.Region=='Japan' ) %>%
  filter( Country.Region %in% top_10_countries) %>%
  mutate( rate = cases  / population  ) %>%
  ggplot(aes( week, rate, color=Country.Region)) +
  geom_line() +
  scale_y_log10() +
  scale_x_continuous()


```


# New cases population rate by day, smoothed


```{r}

coronavirus %>% 
  inner_join(pop, by='name' ) %>% 
  filter( type=='confirmed' ) %>%
  group_by( Country.Region ) %>%
  group_by( Country.Region, date ) %>%
  summarise( cases=sum(cases), population = min(`2015`)) %>%
  ungroup() %>%
  group_by( Country.Region) %>%
  # filter( Country.Region=='Japan' ) %>%
  filter( Country.Region %in% top_10_countries) %>%
  mutate( rate = cases  / population  ) %>%
  ggplot(aes( date, rate, color=Country.Region)) +
  geom_smooth() +
  scale_y_log10() 

```


# # new cases in latest data


```{r}
coronavirus %>% 
  group_by( Country.Region, type) %>%
  filter(date == max(date)) %>%
  ungroup() %>%
  select(country = Country.Region, type, cases, date) %>%
  group_by(country, type) %>%
  summarise(total_cases = sum(cases), date=max(date)) %>%
  pivot_wider(names_from = type, values_from = total_cases) %>%
  arrange(-confirmed)



```

# Australia's choice

We can either follow the slope of Iran/Italy, or China/Korea
Currently, looking like Iran/Italy

```{r selected_countries}

coronavirus %>% 
  #  filter( str_detect(Country.Region, 'ai')) %>%
  filter( Country.Region %in% c('Australia','Italy','Iran','US', 'China', 'Korea, South', 'Taiwan*', 'Japan')) %>%
  inner_join(pop, by='name' ) %>% 
  filter( type=='confirmed' ) %>%
  group_by( Country.Region ) %>%
  mutate(ndays = date-min(date) ) %>%
  group_by( Country.Region, ndays ) %>%
  summarise( cases=sum(cases), population = min(`2015`)) %>%
  ungroup() %>%
  group_by( Country.Region) %>%
  filter( Country.Region %in% c('Australia','Italy','Iran','US', 'China', 'Korea, South', 'Taiwan*', 'Japan')) %>%
  arrange( ndays, .by_group=TRUE) %>%
  mutate( cumcases = cumsum(cases), rate = cumsum(cases ) / population  ) %>%
  ggplot(aes( ndays, rate, color=Country.Region)) +
  geom_line() +
  scale_y_log10() +
  scale_x_continuous()


```

```{r risk}

coronavirus %>% 
filter( Country.Region %in% c('Australia')) %>%
filter(Province.State=='Victoria') %>%
filter( type=='confirmed') %>%
summarise(sum(cases))


```

