---
title:           Covid-19 
author:          Dennis Wollersheim 
date:            06.03.2020
fig_width:       8 
fig_height:      4 
linkcolor:       cyan
citecolor:       grey
output:
  pdf_document:
    highlight:   zenburn
---

\tableofcontents

```{r loadd, results='asis', echo=FALSE, include=FALSE,}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

#devtools::install_github("RamiKrispin/coronavirus")
#install.packages('wpp2017')



```

# Data Source

Sourced from https://ramikrispin.github.io/coronavirus/ and the R package wpp2017, world population profile

If you want to know why Italy just quarantined 18 million people, read on.


```{r Setup}

suppressMessages(suppressWarnings( library(tidyverse)))
library(coronavirus)
library(wpp2017)
data(pop)

pop %>% 
  as_tibble() %>% 
  { . } -> pop

mapping = tribble(
                  ~Country.Region, ~name, 
                  'US', 'United States of America',
                  'UK', 'United Kingdom',
                  'Mainland China', 'China',
                  'South Korea', 'Republic of Korea',
                  'Taiwan','China, Taiwan Province of China',
                  'Vietnam', 'Viet Nam',
                  'Hong Kong', 'China, Hong Kong SAR',
                  'Russia', 'Russian Federation',
                  'Iran','Iran (Islamic Republic of)'
)


coronavirus %>%
  left_join( mapping ) %>% 
  { . } -> coronavirus

```
# Cumulative cases by country and number of days since first case

```{r}


coronavirus %>% 
  filter( type=='confirmed' ) %>%
  group_by( Country.Region ) %>%
  mutate(ndays = date-min(date) ) %>%
  group_by( Country.Region, ndays ) %>%
  summarise( cases=sum(cases)) %>%
  ungroup() %>%
  group_by( Country.Region) %>%
  filter( sum(cases) > 200 ) %>%
  arrange( ndays, .by_group=TRUE) %>%
  mutate( cum_cases = cumsum( cases) ) %>%
  ggplot(aes( ndays, cum_cases, color=Country.Region)) +
  geom_line() +
  scale_y_log10() +
  scale_x_continuous()




```


# Cumulative deaths by country and number of days since first case

```{r}

coronavirus %>% 
  filter( type=='death' ) %>%
  group_by( Country.Region ) %>%
  mutate(ndays = date-min(date) ) %>%
  group_by( Country.Region, ndays ) %>%
  summarise( cases=sum(cases)) %>%
  ungroup() %>%
  group_by( Country.Region) %>%
  filter( sum(cases) > 10 ) %>%
  arrange( ndays, .by_group=TRUE) %>%
  mutate( cum_cases = cumsum( cases) ) %>%
  ggplot(aes( ndays, cum_cases, color=Country.Region)) +
  geom_line() + 
  scale_y_log10() +
  scale_x_continuous()




```

# Rate of new cases per day, since the start of infection by country

 - public health works
 - individualism not so much
 - the hockeys stick curve is worrying people


```{r}

coronavirus %>% 
  filter( type=='confirmed' ) %>%
  filter( Country.Region != 'Others') %>%
  group_by( Country.Region ) %>%
  mutate(ndays = date-min(date) ) %>%
  group_by( Country.Region, ndays ) %>%
  summarise( cases=sum(cases)) %>%
  ungroup() %>%
  group_by( Country.Region) %>%
  filter( sum(cases) > 100 ) %>%
  arrange( ndays, .by_group=TRUE) %>%
  mutate( rate = cumsum(cases ) / as.integer(ndays)  ) %>%
  ggplot(aes( ndays, rate, color=Country.Region)) +
  geom_smooth() +
  scale_y_log10() +
  scale_x_continuous()

```



# Total cumulative confirmed cases as a population rate

Note, Japan had -1 cases on day 2

```{r}

coronavirus %>% 
  mutate( name = ifelse(is.na(name), Country.Region, name)) %>%
  inner_join(pop, by='name' ) %>% 
  filter( type=='confirmed' ) %>%
  group_by( Country.Region ) %>%
  mutate(ndays = date-min(date) ) %>%
  group_by( Country.Region, ndays ) %>%
  summarise( cases=sum(cases), population = min(`2015`)) %>%
  ungroup() %>%
  group_by( Country.Region) %>%
 # filter( Country.Region=='Japan' ) %>%
  filter( sum(cases) > 200 ) %>%
  arrange( ndays, .by_group=TRUE) %>%
  mutate( cumcases = cumsum(cases), rate = cumsum(cases ) / population  ) %>%
  ggplot(aes( ndays, rate, color=Country.Region)) +
  geom_line() +
  scale_y_log10() +
  scale_x_continuous()

```


# # new cases in latest data


```{r}
coronavirus %>% 
  group_by( Country.Region, type) %>%
  filter(date == max(date)) %>%
  ungroup() %>%
  select(country = Country.Region, type, cases, date) %>%
  group_by(country, type) %>%
  summarise(total_cases = sum(cases), date=max(date)) %>%
  pivot_wider(names_from = type, values_from = total_cases) %>%
  arrange(-confirmed)



```
